---
title: "DS_Session_4"
output: html_document
author: Suchir Kalra
date: "`r Sys.Date()`"
fontsize: 12pt

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the libraries and reading the dataset 

```{r, include=FALSE}
library(tidyverse)
library(knitr) # Load the 'knitr' package, which is useful for dynamic report generation in R
library(ggthemes)# Load the 'ggthemes' package, which provides additional themes for ggplot2 visualizations
library (ggplot2)

tcpd = read.csv('All_States_GE.csv')
```

## Looking at what kable() does in the RMarkdown file! 

```{r Delhi Turnout}
Delhi = tcpd %>% 
  filter(State_Name == 'Delhi' & Year >= 2004 & Year <=2019 & Poll_No == 0)

Delhi_turnout = Delhi %>% 
  filter(Position==1)%>%
  group_by(Year, Constituency_Name)%>%
  summarise(turnout=mean(Turnout_Percentage,na.rm=T)) %>% 
  mutate (State = "Delhi")

Delhi_turnout %>% 
  kable()

```

Now for Haryana Turnout

```{r Haryana Turnout}

Haryana = tcpd %>% 
  filter(State_Name == 'Haryana' & Year >= 2004 & Year <=2019 & Poll_No == 0)

Haryana_turnout = Haryana %>% 
  filter(Position==1)%>%
  group_by(Year, Constituency_Name)%>%
  summarise(turnout=mean(Turnout_Percentage,na.rm=T)) %>% 
  mutate (State = "Haryana")


Haryana_turnout %>% 
  kable()
```

Pretty decent! It makes a table for you that can be used in reports! Turnout in Delhi has gone up in 2014 elections...

## GGPLOT! Of Delhi & Haryana Turnout! BoxPlot!


```{r plotting of Delhi & Haryana Turnout, echo=FALSE} 

##If you’re preparing a report or a presentation, you might want to show results (tables, graphs) without cluttering the document with R code. Thus, echo = FALSE -- this makes sure that the RCode is not included in the report)

combined_turnout = bind_rows(Delhi_turnout, Haryana_turnout) 

# Create faceted box plots
ggplot(combined_turnout, aes(x = factor(Year), y = turnout)) +
  geom_boxplot() +
  facet_wrap(~State) +  # Creates separate panels for Delhi & Haryana
  labs(title = "Voter Turnout in Delhi and Haryana (2004-2019)", 
       x = "Year", 
       y = "Turnout Percentage") +
  theme_minimal()


```

Analysis: Turnout overall higher in Haryana because of door-to-door panchayats pulling out voters to go and vote? 

Conclusion: Can add conclusion or initial observations here..

Note: A single # as below is for title and double ## is for subtitle etc.

# GGPLOT DS 4 Practice

## Exercise 1: Change in Directly Recruited Women in IAS (1979-2015)

### Descriptives on Women in IAS 


```{r ias-gender-over-time, echo=FALSE, message=FALSE}
#############################################

ias.profile = read.csv("ias-profile.csv") 

##Aggregate by Year and Allotment and Gender 
##Tells you total number of directly recruited women every year 
overtime_gender<-ias.profile%>%
  filter(Source_of_Recruitment=="Direct Recruitment")%>%
  filter(Allotment_Year>1979&Allotment_Year<2015)%>%
  group_by(Allotment_Year,Gender)%>%
  summarise(Number=n())%>%
  ungroup() 
#############################################
###Create a total number of recruited value 
###Gender M/F both rows and being added 
overtime_gender<-overtime_gender%>%
  group_by(Allotment_Year)%>%
  mutate(Total=sum(Number))%>%
  ungroup()

#############################################
#Create Percentage Women Variable 
overtime_gender$Percentage<-(overtime_gender$Number/overtime_gender$Total)*100

#############################################
#Map the change in % women over time 
ggplot(overtime_gender, aes(x=Allotment_Year, y=Percentage, fill=Gender )) +
  geom_line(aes(color=Gender))+ggtitle("Change in Gender Representation in IAS (Direct) Over Time")+geom_point()+xlab("Year")
```

## Exercise 2: State-wise turnout in the 2019 elections 


```{r State_Wise Turnout in the 2019 elections}

State_turnout = tcpd %>% 
  filter(Year == 2019 & Poll_No == 0 & Position==1) %>% 
  group_by(State_Name)%>%
  summarise(turnout=mean(Turnout_Percentage,na.rm=T))

State_turnout %>% 
  kable()
```
## Make a Graph of the State-wise turnout 

```{r State_Wise_Turnout_2019, echo=FALSE, message=FALSE, warning=FALSE}

library(ggplot2)

# Create the bar plot
ggplot(State_turnout, aes(x = reorder(State_Name, -turnout), y = turnout)) +
  geom_col(fill = "steelblue") +
  coord_flip() +  # Flip to make it easier to read
  labs(title = "State-Wise Voter Turnout in 2019 Elections",
       x = "State",
       y = "Average Turnout Percentage") +
  theme_minimal()

#Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```


## Now, do it for North, South, West and East India

```{r Clean_State_Names, echo=FALSE, message=FALSE, warning=FALSE}
# Replace underscores with spaces in the State_Name column
State_turnout <- State_turnout %>%
  mutate(State_Name = gsub("_", " ", State_Name))
```


```{r Region_Wise_Turnout_2019, echo=FALSE, message=FALSE, warning=FALSE}

# Define the regional mapping
north_states <- c("Jammu & Kashmir","Himachal Pradesh", "Punjab", "Uttarakhand", 
                  "Haryana", "Uttar Pradesh")

south_states <- c("Andhra Pradesh", "Karnataka", "Kerala", "Tamil Nadu", 
                  "Telangana")

east_states <- c("Bihar", "Jharkhand", "Odisha", "West Bengal", "Sikkim", 
                 "Assam", "Arunachal Pradesh", "Meghalaya", "Manipur", 
                 "Mizoram", "Nagaland", "Tripura")

west_states <- c("Goa", "Gujarat", "Maharashtra", "Rajasthan")

central_states <- c("Madhya Pradesh", "Chhattisgarh")

# Assign states to regions
State_turnout <- State_turnout %>%
  mutate(Region = case_when(
    State_Name %in% north_states ~ "North",
    State_Name %in% south_states ~ "South",
    State_Name %in% east_states ~ "East",
    State_Name %in% west_states ~ "West",
    State_Name %in% central_states ~ "Central",
    TRUE ~ "UT"
  ))

# Aggregate turnout by region
Region_turnout <- State_turnout %>%
  group_by(Region) %>%
  summarise(Avg_Turnout = mean(turnout, na.rm = TRUE))

# Plot the region-wise turnout
ggplot(Region_turnout, aes(x = Region, y = Avg_Turnout, fill = Region)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Region-Wise Voter Turnout in 2019 Elections",
       x = "Region",
       y = "Average Turnout Percentage") +
  theme_minimal()
```


