---
title: "DS_Session_4"
output: html_document
author: Suchir Kalra
date: "`r Sys.Date()`"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the libraries and reading the dataset 

```{r, include=FALSE}
library(tidyverse)
library(knitr) # Load the 'knitr' package, which is useful for dynamic report generation in R
library(ggthemes)# Load the 'ggthemes' package, which provides additional themes for ggplot2 visualizations
library (ggplot2)

tcpd = read.csv('All_States_GE.csv')
```

## Looking at what kable() does in the RMarkdown file! 

```{r Delhi Turnout}
Delhi = tcpd %>% 
  filter(State_Name == 'Delhi' & Year >= 2004 & Year <=2019 & Poll_No == 0)

Delhi_turnout = Delhi %>% 
  filter(Position==1)%>%
  group_by(Year, Constituency_Name)%>%
  summarise(turnout=mean(Turnout_Percentage,na.rm=T)) %>% 
  mutate (State = "Delhi")

Delhi_turnout %>% 
  kable()

```

Now for Haryana Turnout

```{r Haryana Turnout}

Haryana = tcpd %>% 
  filter(State_Name == 'Haryana' & Year >= 2004 & Year <=2019 & Poll_No == 0)

Haryana_turnout = Haryana %>% 
  filter(Position==1)%>%
  group_by(Year, Constituency_Name)%>%
  summarise(turnout=mean(Turnout_Percentage,na.rm=T)) %>% 
  mutate (State = "Haryana")


Haryana_turnout %>% 
  kable()
```

Pretty decent! It makes a table for you that can be used in reports!

## GGPLOT! Of Delhi & Haryana Turnout! BoxPlot!


```{r plotting of Delhi & Haryana Turnout , echo=FALSE} 

##If you’re preparing a report or a presentation, you might want to show results (tables, graphs) without cluttering the document with R code. Thus, echo = FALSE -- this makes sure that the RCode is not included in the report)

combined_turnout = bind_rows(Delhi_turnout, Haryana_turnout) 

# Create faceted box plots
ggplot(combined_turnout, aes(x = factor(Year), y = turnout)) +
  geom_boxplot() +
  facet_wrap(~State) +  # Creates separate panels for Delhi & Haryana
  labs(title = "Voter Turnout in Delhi and Haryana (2004-2019)", 
       x = "Year", 
       y = "Turnout Percentage") +
  theme_minimal()


```

# GGPLOT DS 4 Practice
## Exercise 1: Change in Directly Recruited Women in IAS (1979-2015)
### Descriptives on Women in IAS 

A single # as above is for title and double ## is for subtitle etc. 

```{r ias-gender-over-time, echo=FALSE, message=FALSE}
#############################################

ias.profile = read.csv("ias-profile.csv") 

##Aggregate by Year and Allotment and Gender 
##Tells you total number of directly recruited women every year 
overtime_gender<-ias.profile%>%
  filter(Source_of_Recruitment=='Direct Recruitment')%>%
  filter(Allotment_Year>1979&Allotment_Year<2015)%>%
  group_by(Allotment_Year,Gender)%>%
  summarise(Number=n())%>%
  ungroup()
#############################################
###Create a total number of recruited value 
###Gender M/F both rows and being added 
overtime_gender<-overtime_gender%>%
  group_by(Allotment_Year)%>%
  mutate(Total=sum(Number))%>%
  ungroup()

#############################################
#Create Percentage Women Variable 
overtime_gender$Percentage<-overtime_gender$Number/overtime_gender$Total

#############################################
#Map the change in % women over time 
ggplot(overtime_gender, aes(x=Allotment_Year, y=Percentage, fill=Gender )) +
  geom_line(aes(color=Gender))+ggtitle("Change in Gender Representation in IAS (Direct) Over Time")+geom_point()+xlab("Year")
```



